---
title: "Data Science Interview Project"
author: "Jordan Cochuyt"
date: "2023-08-29"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(tidymodels)
library(summarytools)
library(gtsummary)
library(arsenal)
library(randomForest)
library(caret)
library(pROC)
library(ROCR)
library(Hmisc)
library(ROSE)
library(DT)

## Pulling in the text files with the data
claims_test <- read.table(file = "Claims_Test.txt", header = TRUE)                          ## Claims Dataset used for Testing the Model
claims_train <- read.table(file = "Claims_Train.txt", header = TRUE)                        ## Claims Dataset used for Training the Model
location_of_service <- read.table(file = "Location_Of_Service.txt", header = TRUE)          ## Information for location of Services | Will use in case logic
surgeries_test <- read.table(file = "Surgeries_Test.txt", header = TRUE)                    ## Surgeries Dataset used for Testing the Model
surgeries_train <- read.table(file = "Surgeries_Train.txt", header = TRUE)                  ## Surgeries Dataset used for Training the Model

############################### Training Data Setup ######################################################
claims_train <- claims_train %>% mutate(LOS_DESC = as.factor(case_when(LOS_CODE == 23 ~ "Emergency Room",                                 ## Linking the location of service categories into a new variable 
                              LOS_CODE == 22 ~ "Outpatient",
                              LOS_CODE == 21 ~ "Inpatient",
                              LOS_CODE == 11 ~ "Professional",
                              LOS_CODE == 20 ~ "Urgent Care",
                              LOS_CODE == 34 ~ "Hospice",
                              LOS_CODE == 2 ~ "Telehealth")))

#### Making a flag for Outpatient LOS
outpatient_ids <- claims_train %>% filter(LOS_DESC == "Outpatient") %>% select(MEMBER_ID)
outpatient_ids <- outpatient_ids[,1]

#### Professional flag ids
professional_ids <- claims_train %>% filter(LOS_DESC == "Professional") %>% select(MEMBER_ID)
professional_ids <- professional_ids[,1]

#### Inpatient flag ids
inpatient_ids <- claims_train %>% filter(LOS_DESC  == "Inpatient") %>% select(MEMBER_ID)
inpatient_ids <- inpatient_ids[,1]

### Number of claims
claims_num <- claims_train %>% group_by(MEMBER_ID) %>% summarise(claim_count = n())

### New claims training dataset
claims_num <- claims_num %>% mutate(outpatient = case_when(MEMBER_ID %in% outpatient_ids ~ "Yes",
                                                           TRUE ~ "No"),
                                    professional = case_when(MEMBER_ID %in% professional_ids ~ "Yes",
                                                              TRUE ~ "No "),
                                    inpatient = case_when(MEMBER_ID %in% inpatient_ids ~ "Yes",
                                                           TRUE ~ "No"))

training.data <- left_join(surgeries_train, claims_num, by = "MEMBER_ID") %>%                  ## Joining the training datasets together by MEMBER_ID
  mutate(
         ### Making outcome SURGERY a Factor (categorical variable)
         SURGERY = as.factor(case_when(SURGERY == "no" ~ "No",
                                       SURGERY == "yes" ~ "Yes")),
         ### Making DIABETES a Factor (categorical variable)
         DIABETES = as.factor(case_when(DIABETES == 0 ~ "No",
                              DIABETES == 1 ~ "Yes")),
         ### Making OPIOID_RX a Factor (categorical variable)
         OPIOID_RX = as.factor(case_when(OPIOID_RX == 0 ~ "No",
                                         OPIOID_RX == 1 ~ "Yes")),
         ### Making CANCER a Factor (categorical variable)
         CANCER = as.factor(case_when(CANCER == 0 ~ "No",
                                      CANCER == 1 ~ "Yes")),
         ### Making Gender a Factor (categorical variable)
         GENDER = as.factor(GENDER)
         )

### Imputing the median value for claims_count
training.data <- training.data %>% mutate(claim_count = impute(claim_count, median))

##################################### Testing Data Setup ################################################
claims_test <- claims_test %>% mutate(LOS_DESC = as.factor(case_when(LOS_CODE == 23 ~ "Emergency Room",                                 ## Linking the location of service categories into a new variable 
                              LOS_CODE == 22 ~ "Outpatient",
                              LOS_CODE == 21 ~ "Inpatient",
                              LOS_CODE == 11 ~ "Professional",
                              LOS_CODE == 20 ~ "Urgent Care",
                              LOS_CODE == 34 ~ "Hospice",
                              LOS_CODE == 2 ~ "Telehealth")))

#### Making a flag for Outpatient LOS
outpatient_ids_test <- claims_test %>% filter(LOS_DESC == "Outpatient") %>% select(MEMBER_ID)
outpatient_ids_test <- outpatient_ids_test[,1]

#### Professional flag ids
professional_ids_test <- claims_test %>% filter(LOS_DESC == "Professional") %>% select(MEMBER_ID)
professional_ids_test <- professional_ids_test[,1]

#### Inpatient flag ids
inpatient_ids_test <- claims_test %>% filter(LOS_DESC  == "Inpatient") %>% select(MEMBER_ID)
inpatient_ids_test <- inpatient_ids_test[,1]

### Number of claims
claims_num_test <- claims_test %>% group_by(MEMBER_ID) %>% summarise(claim_count = n())

### New claims training dataset
claims_num_test <- claims_num_test %>% mutate(outpatient = case_when(MEMBER_ID %in% outpatient_ids_test ~ "Yes",
                                                           TRUE ~ "No"),
                                    professional = case_when(MEMBER_ID %in% professional_ids_test ~ "Yes",
                                                              TRUE ~ "No "),
                                    inpatient = case_when(MEMBER_ID %in% inpatient_ids_test ~ "Yes",
                                                           TRUE ~ "No"))

testing.data <- left_join(surgeries_test, claims_num_test, by = "MEMBER_ID") %>%                  ## Joining the training datasets together by MEMBER_ID
  mutate(
         ### Making DIABETES a Factor (categorical variable)
         DIABETES = as.factor(case_when(DIABETES == 0 ~ "No",
                              DIABETES == 1 ~ "Yes")),
         ### Making OPIOID_RX a Factor (categorical variable)
         OPIOID_RX = as.factor(case_when(OPIOID_RX == 0 ~ "No",
                                         OPIOID_RX == 1 ~ "Yes")),
         ### Making CANCER a Factor (categorical variable)
         CANCER = as.factor(case_when(CANCER == 0 ~ "No",
                                      CANCER == 1 ~ "Yes")),
         ### Making Gender a Factor (categorical variable)
         GENDER = as.factor(GENDER)
         )

### Imputing the missing claim counts to the median of the training to stay consitent between the two
testing.data <- testing.data %>% mutate(claim_count = impute(claim_count, 2))

```

# Exploritory Data Analysis
## Descriptive Statistics of the Training Data
```{r summary}
### Taking out any variables that do not need a summary i.e. MEMBER_ID
dfsummary.data <- training.data %>% select(SURGERY, WEIGHTED_RISK_SCORE, DIABETES, OPIOID_RX, CANCER, AGE, GENDER, claim_count, outpatient, professional, inpatient)

### Table displaying summary statistics from the summarytools package
print(dfSummary(dfsummary.data,
          varnumbers = FALSE, 
          valid.col = FALSE, 
          graph.magnif = 0.76),
      method = 'render')
```

## Descriptive Statistics of the Training Data by SURGERY
```{r, results="asis"}
tab.surg <- tableby(SURGERY ~ ., data = dfsummary.data)
summary(tab.surg, text = TRUE) 
```

## Box Plots 

### Inpatient vs Non-Inpatient Claim Count

```{r, warning=FALSE} 
ggplot(aes(x = inpatient, y = claim_count), data = dfsummary.data) + geom_boxplot() + xlab("Inpatient") + ylab("# of Claims") + theme_classic() + theme(legend.position = "none")
```

### Professional vs Non-Professional Claim Count

```{r, warining = FALSE}
ggplot(aes(x = professional, y = claim_count), data = dfsummary.data) + geom_boxplot() + xlab("Professional") + ylab("# of Claims") + theme_classic() + theme(legend.position = "none")
```

### Outpatient vs Non-Outpatient Claim Count

```{r} 
ggplot(aes(x = outpatient, y = claim_count), data = dfsummary.data) + geom_boxplot() + xlab("Outpatient") + ylab("# of Claims") + theme_classic() + theme(legend.position = "none")
```

### DIABETES vs Non-Diabetes Claim Count

```{r}
ggplot(aes(x = DIABETES, y = claim_count), data = dfsummary.data) + geom_boxplot() + xlab("Diabetes") + ylab("# of Claims") + theme_classic() + theme(legend.position = "none")
```

### DIABETES vs Non-Diabetes Weighted Risk Score

```{r}
ggplot(aes(x = DIABETES, y = WEIGHTED_RISK_SCORE), data = dfsummary.data) + geom_boxplot() + xlab("Diabetes") + ylab("Risk Score") + theme_classic() + theme(legend.position = "none")
```

### DIABETES vs Non-Diabetes  Age

```{r}
ggplot(aes(x = DIABETES, y = AGE), data = dfsummary.data) + geom_boxplot() + xlab("Diabetes") + ylab("Age") + theme_classic() + theme(legend.position = "none")
```

### OPIOIDs Claim Count

```{r}
ggplot(aes(x = OPIOID_RX, y = claim_count), data = dfsummary.data) + geom_boxplot() + xlab("Opioid Use") + ylab("# of Claims") + theme_classic() + theme(legend.position = "none")
```

### OPIOIDs Weighted Risk Score

```{r}
ggplot(aes(x = OPIOID_RX, y = WEIGHTED_RISK_SCORE), data = dfsummary.data) + geom_boxplot() + xlab("Opioid Use") + ylab("Risk Score") + theme_classic() + theme(legend.position = "none")
```

### OPIOIDs Age

```{r}
ggplot(aes(x = OPIOID_RX, y = AGE), data = dfsummary.data) + geom_boxplot() + xlab("Opioid Use") + ylab("Age") + theme_classic() + theme(legend.position = "none")
```

## Univariate Logistic Regression to check for association
```{r}
### Univariate Logistic Regression to check for associations between the covariates and our surgery outcome
dfsummary.data %>%
  tbl_uvregression(method = glm,
                   y = factor(SURGERY),
                   method.args = list(family = binomial),
                   exponentiate = TRUE
                   )

```

## Multivariate Logistic Regression Model
### Variable Selection for Model
1) Removing the key id variables for the individuals (i.e. MEMBER_ID, CLAIM_ID)  
2) Remove ICD10: There are simply to many categories currently for this to be in the model. An interesting piece that could be added here would be to include area of diagnosis. This may show areas that are more likely than others to receive a surgery and help strengthen a prediction.    
3) Remove Cancer: The reason for removing cancer we simply do not have enough individuals with surgery to use it in the model (count of 5). For this reason I am pulling the variable from the data set.  
4) Remove Gender: When we look above at the univariate logistic model gender is not significant (p-value: .4), for a significant association between having a surgery and the gender of an individual. Since there is no significant evidence here I will not include it in the model.  
5) Remove Diabetes: When we look above at the univariate logistic model Diabetes is not significant (p-value: .8).  

```{r,}

### Setting the randomization seed for reproducibility
set.seed(12345)

### Pulling out some data to validate the model and see its performance
training.final.split <- initial_split(dfsummary.data, prop = .75,
                                       strata = SURGERY)

training.train <- training.final.split %>% training()
training.test <- training.final.split %>% testing()

### Specifying the model type
logistic_model <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

### Specifying the recipe
  logistic_rec <- recipe(SURGERY ~ WEIGHTED_RISK_SCORE + AGE + OPIOID_RX + claim_count, data = training.train)
  
### Making the workflow
  logistic_workflow <- workflow() %>%
    add_model(logistic_model) %>%
    add_recipe(logistic_rec)
  
### Fit the model to the data
  logistic_model_fit <- logistic_workflow %>%
    fit(data = training.train)

tidy(logistic_model_fit, exponentiate = TRUE)
```

## Model Prediction using Testing Data
```{r}
## Run validation of the model
logistic_prob_preds <- predict(logistic_model_fit,
                new_data = training.test,
                type = "prob")

logistic_results <- training.test %>% 
  select(SURGERY) %>%
  bind_cols(logistic_prob_preds)

### Pulling the top 6% of the pred_Yes to be classed as a "Yes" for surgery
top2.logistic.preds <- logistic_results %>% slice_max(order_by=.pred_Yes, prop = 0.02) %>% mutate(pred.class = "Yes")

### Pulling the bottom 94% of the pred_Yes to be calssed as a "No" for surgery
bottom98.logistic.preds <- logistic_results %>% slice_min(order_by = .pred_Yes, prop = 0.98) %>% mutate(pred.class = "No")

logistic.pred.final <- rbind(top2.logistic.preds, bottom98.logistic.preds) %>% mutate(pred.class = as.factor(pred.class))

```

### Confidence Matrix
```{r}
conf_mat(logistic.pred.final, truth = SURGERY,
         estimate = pred.class)
```

### Accuracy
```{r}
accuracy(logistic.pred.final, truth = SURGERY,
         estimate = pred.class)
```

#### ROC CURVE
```{r}
roc_curve(logistic.pred.final, truth = SURGERY, .pred_Yes, event_level = "second") %>% autoplot()
```

### AUC
```{r}
roc_auc(logistic.pred.final, truth = SURGERY, .pred_Yes, event_level = "second")
```

## Random Forest Full Model (No-resample)

### Random Forest Model
```{r}
randomforest.data <- training.train %>% select(-outpatient, -inpatient, -professional)

rf <- randomForest(SURGERY ~ ., data = randomforest.data,  mtry = 4, ntree = 250, importance = TRUE) ## 4 Levels to decision trees, 1000 trees for each prediction, keep the importance metrics for variables

rf

```

#### Importance of Variables

```{r}
importance = importance(rf)
varImportance = data.frame(Variables = row.names(importance),
                           Importance = round(importance[, "MeanDecreaseAccuracy"], 2))

rankImportance = varImportance %>% mutate(Rank = paste('#', dense_rank(desc(Importance))))

ggplot(rankImportance, aes(x = reorder(Variables,Importance), y= Importance, fill = Importance))  + 
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'white') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_classic()
```

#### Validation of Full Model
```{r}
randomforest.test <- training.test %>% filter(is.na(outpatient))
retain_p <- rf %>% predict(newdata = training.test)

table(
  actualclass = training.test$SURGERY,
  predictedclass = retain_p
) %>% 
  confusionMatrix(positive='Yes') %>% 
  print()
```

### Reduced Random Forest Model
```{r}
randomforest.data.2 <- randomforest.data %>% select(-CANCER, -GENDER, - DIABETES)
rf2 <- randomForest(SURGERY ~ ., data = randomforest.data.2,  mtry = 4, ntree = 250, importance = TRUE) ## 4 Levels to decision trees, 1000 trees for each prediction, keep the importance metrics for variables

rf2

```

#### Importance of Variables

```{r}
importance = importance(rf2)
varImportance = data.frame(Variables = row.names(importance),
                           Importance = round(importance[, "MeanDecreaseAccuracy"], 2))

rankImportance = varImportance %>% mutate(Rank = paste('#', dense_rank(desc(Importance))))

ggplot(rankImportance, aes(x = reorder(Variables,Importance), y= Importance, fill = Importance))  + 
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'white') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_classic()
```

#### Validation of Reduced Model
```{r}
randomforest.test <- training.test %>% filter(is.na(outpatient))
retain_p <- rf2 %>% predict(newdata = training.test)

table(
  actualclass = training.test$SURGERY,
  predictedclass = retain_p
) %>% 
  confusionMatrix(positive='Yes') %>% 
  print()
```

## Random Forest Data Resampling (Over-Sampling)
```{r}
### Resampling the Data to get rid of the imbalanced outcome
randomforest.data.resample <- ovun.sample(SURGERY ~ ., data = randomforest.data, method = "over", N = 50000, seed = 1234)$data

freq(randomforest.data.resample$SURGERY)

```

### Resampled Random Forest Full Model
```{r}
### Random Forest Model
rf3 <- randomForest(SURGERY ~ ., data = randomforest.data.resample,  mtry = 5, ntree = 250, importance = TRUE) ## 4 Levels to decision trees, 1000 trees for each prediction, keep the importance metrics for variables

rf3

```

#### Variable Importance
```{r}
importance = importance(rf3)
varImportance = data.frame(Variables = row.names(importance),
                           Importance = round(importance[, "MeanDecreaseAccuracy"], 2))

rankImportance = varImportance %>% mutate(Rank = paste('#', dense_rank(desc(Importance))))

ggplot(rankImportance, aes(x = reorder(Variables,Importance), y= Importance, fill = Importance))  + 
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'white') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_classic()
```

#### Validation of Full Model
```{r}
retain_p <- rf3 %>% predict(newdata = training.test)

table(
  actualclass = training.test$SURGERY,
  predictedclass = retain_p
) %>% 
  confusionMatrix(positive='Yes') %>% 
  print()
```

### Model choice for official predictions: Reduced Random Forest Model
```{r}
model.testing <- rf2 %>% predict(newdata = testing.data, "prob")

predictions <- data.frame(MEMBER_ID = testing.data$MEMBER_ID, Prediction = model.testing[,2])

write.csv(predictions, file = "C:\\Users\\jorda\\OneDrive\\Desktop\\predictions.csv")
### Threshold will be 0.408, because the top 2% of the predictions ends at that probability. 2% of Training data had surgery so we want to reflect that.
```

## Final Product
#### Threshold: 0.408
I have chosen to use the threshold 0.408 for my prediction probability classification. The reason for this being that is the threshold that reflects the top 2% of the predictions, which was the same as the percentage of surgeries within the training sample. The thought behind this is if we are predicting surgeries 2% of the testing data should be classified as surgeries. 

#### What surgery are we predicting?
#### Top 50 ICD-10 Procedure Codes for Surgery Claims in the Dataset
```{r}
train_desc_surg <- training.data %>% filter(SURGERY == "Yes")                ### Limit the data to only individuals that underwent surgery
ICD10.codes <- freq(train_desc_surg$ICD10, order = "freq") %>% tb()          ### Get the counts of the individual ICD-10 codes
top50.codes <- ICD10.codes[1:50,1:2]                                         ### Keep the top 50 codes and the frequencies associated

### Manually insert the ICD-10 Code Descriptions
top50 <- top50.codes %>% mutate(CODE_DESCRIPTION = case_when(ICD10 == "M79651" ~ "Pain in right thigh",
                                                             ICD10 == "M25529" ~ "Pain in unspecified elbow",
                                                             ICD10 == "M79674" ~ "Pain in right toe(s)",
                                                             ICD10 == "M79642" ~ "Pain in left hand",
                                                             ICD10 == "M25551" ~ "Pain in right hip",
                                                             ICD10 == "M25562" ~ "pain experienced in the left knee",
                                                             ICD10 == "M7910" ~ "Fistula, unspecified joint",
                                                             ICD10 == "M25659" ~ "Stiffness of unspecified hip, not elsewhere classified",
                                                             ICD10 == "M79644" ~ "Pain in right finger(s)",
                                                             ICD10 == "M25641" ~ "Stiffness of right hand, not elsewhere classified",
                                                             ICD10 == "M25652" ~ "Stiffness of left hip, not elsewhere classified",
                                                             ICD10 == "M62830" ~ "back muscle spasms",
                                                             ICD10 == "M25561" ~ "pain in the right knee",
                                                             ICD10 == "M25651" ~ "Stiffness of right hip, not elsewhere classified",
                                                             ICD10 == "M79652" ~ "Pain in left thigh",
                                                             ICD10 == "M25675" ~ "Stiffness of left foot, not elsewhere classified",
                                                             ICD10 == "M79639" ~ "Pain in unspecified forearm",
                                                             ICD10 == "M79659" ~ "Pain in unspecified thigh",
                                                             ICD10 == "M79669" ~ "Pain in unspecified lower leg",
                                                             ICD10 == "M25631" ~ "Stiffness of right wrist, not elsewhere classified",
                                                             ICD10 == "M79602" ~ "Pain in left arm",
                                                             ICD10 == "M25542" ~ "Pain in joints of left hand",
                                                             ICD10 == "M25552" ~ "Pain in Left Hip",
                                                             ICD10 == "M25611" ~ "Stiffness of right shoulder, not elsewhere classified",
                                                             ICD10 == "M25579" ~ "Pain in unspecified ankle and joints of unspecified foot",
                                                             ICD10 == "M2550" ~ "Pain in the Unspecified Joint",
                                                             ICD10 == "M79643" ~ "Pain in unspecified hand",
                                                             ICD10 == "M79673" ~ "Pain in unspecified foot",
                                                             ICD10 == "M25612" ~ "Stiffness of left shoulder, not elsewhere classified",
                                                             ICD10 == "M25539" ~ "Pain in unspecified wrist",
                                                             ICD10 == "M25672" ~ "Stiffness of left ankle, not elsewhere classified",
                                                             ICD10 == "M79609" ~ "Pain in unspecified limb",
                                                             ICD10 == "M25559" ~ "Pain in unspecified hip",
                                                             ICD10 == "M25661" ~ "Stiffness of right knee, not elsewhere classified",
                                                             ICD10 == "M5481" ~ "occipital neuralgia",
                                                             ICD10 == "M79629" ~ "Pain in unspecified upper arm",
                                                             ICD10 == "M79671" ~ "pain in the right foot",
                                                             ICD10 == "M25511" ~ "Pain in Right Shoulder",
                                                             ICD10 == "M25572" ~ "Pain experienced specifically in the left ankle and the joints of the left foot",
                                                             ICD10 == "M25639" ~ "Stiffness of unspecified wrist, not elsewhere classified",
                                                             ICD10 == "M25674" ~ "Stiffness of right foot, not elsewhere classified",
                                                             ICD10 == "M71439" ~ "Calcium deposit in bursa, unspecified wrist",
                                                             ICD10 == "M7912" ~ "Myalgia of auxiliary muscles, head and neck",
                                                             ICD10 == "M7918" ~ "Myalgia, other site",
                                                             ICD10 == "M79646" ~ "Pain in unspecified finger(s)",
                                                             ICD10 == "M25519" ~ "pain in unspecified shoulder",
                                                             ICD10 == "M79622" ~ "Pain in left upper arm",
                                                             ICD10 == "M79631" ~ "Pain in right forearm",
                                                             ICD10 == "K4031" ~ "Unilateral inguinal hernia, with obstruction, without gangrene, recurrent",
                                                             ICD10 == "M25531" ~ "Pain in right wrist"))

### Display the table
top50 <- top50 %>% select(ICD10, CODE_DESCRIPTION, freq)

DT::datatable(top50)
```

From looking at the top 50 ICD-10 codes people that had a surgery went through, I believe we are looking at joint procedures for a primary diagnosis involving arthritis. 
